DRAFT: Harmonized and trait aggregation paper

### 2020-10-20 Paper/Draft_harmonization_paper.tex
20.10: Updat Paper

@@

Aggregated trait affinities of the five trait aggregation methods (\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg}) were compared to trait affinities assigned at family-level by experts, which were available for Australia and North America for a subset of grouping features and taxa. For the Australian dataset, we compared aggregated trait affinities with assigned trait affinities resolved at family-level for the grouping features feeding mode and size using data from Chessman \cite{chessman_dissolved-oxygen_2018}. In Chessman \cite{chessman_dissolved-oxygen_2018} feeding mode is classified similarly as in the Australian dataset except that the trait parasite is missing. We conducted the comparison for the 220 families available in Chessman \cite{chessman_dissolved-oxygen_2018}. Considering each factor combination of family and trait (in total 8) this amounts to 1760 cases.

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the North American database (Twardochleb) \cite{twardochleb_trait_data_2020}) and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

% Sampled from a truncated normal distribution values for three traits
% (upper and lower boundaries 0 and 1)
% Traits were normalized so that they sum up to one
% Three different taxonomic hierarchies were simulated: 
% Simulated one family with 25 species 
% -> 1) 5 Genera:  5 Species respectively
% -> 2) 5 Genera: 1 Genus 13 Species (more than ~ 50 %); 4 Genera with 3 species
% -> 3) 5 Genera: All have different amount of species

% Simulation result: 
% -  base simulation with similar amount of speices per genus and genera per family and varying trait values
% results can vary when variation in trait values increases 
% However, mostly the different aggregation methods result in similar trait affinities
% when differences occur, only for: 
% dir_median vs stepwise median
% direct mean vs stepwise median
% stepwise mean vs stepwise median
% direct mean vs direct median
% stepwise median vs weighed agg
% Trait variability: 0.3 and greater
% Max diff ~ 0.2

>>>>

% BK:
% # I wonder if some simulation could be useful to determine in what
% # circumstances the different aggregation methods will produce similar
% # or different results. E.g. If all families were to have the same number of
% # genus and all genus the same number of species (something which would
% #  never happen),
% # then it will not matter (much) whether direct, stepwise or
% # weighted aggregation is used.
% # It only becomes important if direct, stepwise or weighted is used,
% # as the number of genus per family and/or species per genus differ.
% # Some simulation that varies the taxonomic hierarchy and variability in traits
% # between genus within a family and species within genus, ….,

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we simulated three different taxonomic hierarchies. 
Sim1) A family with equal number of genera and species (5 genera each with 5 species respectively).
Sim2) A family were one genus has a much larger number of species as the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively). 
Sim3) A family were all genera have a different number of species (8, 2, 7, 3, 5)
Every simulated taxonomic hierarchy was assigned a hypothetical grouping feature with 3 traits. The individual affinities for each trait were sampled from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviation (0.2, 0.4, 0.6, 0.8 and 1) respectively to simulate trait variability. The truncated normal distribution was bound to 0 and 1. The traits were normalized so that they sum up to one for each species, similar to the data processing of the trait databases. 100 Simulations for each standard deviation were carried out. Hence, 
500 simulations were run for each simulation scenario (in total 1500). 
All trait aggregation methods were applied to each simulated dataset and the results were compared.   
% Sampled from a truncated normal distribution values for three traits
% (upper and lower boundaries 0 and 1)
% Traits were normalized so that they sum up to one
% Three different taxonomic hierarchies were simulated: 
% Simulated one family with 25 species 
% -> 1) 5 Genera:  5 Species respectively
% -> 2) 5 Genera: 1 Genus 13 Species (more than ~ 50 %); 4 Genera with 3 species
% -> 3) 5 Genera: All have different amount of species

@@

Aggregated trait affinities of the five trait aggregation methods (\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg}) were compared to trait affinities assigned at family-level by experts, which were available for Australia and North America for a subset of grouping features and taxa. For the Australian dataset, we compared aggregated trait affinities with assigned trait affinities resolved at family-level for the grouping features feeding mode and size using data from Chessman et al. \cite{chessman_dissolved-oxygen_2018}. In Chessman et al. \cite{chessman_dissolved-oxygen_2018} feeding mode is classified similarly as in the Australian dataset except that the trait parasite is missing. We conducted the comparison for the 220 families available in Chessman et al. \cite{chessman_dissolved-oxygen_2018}. Considering each factor combination of family and trait (in total 8) this amounts to 1760 cases.

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the North American database (Twardochleb et al.) \cite{twardochleb_trait_data_2020} and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

% Simulation result: 
% -  base simulation with similar amount of species per genus and genera per family and varying trait values
% results can vary when variation in trait values increases 
% However, mostly the different aggregation methods result in similar trait affinities
% when differences occur, only for: 
% dir_median vs stepwise median
% direct mean vs stepwise median
% stepwise mean vs stepwise median
% direct mean vs direct median
% stepwise median vs weighed agg
% Trait variability: 0.3 and greater
% Max diff ~ 0.2
